Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-12 BN_discretizer_paper.tex
made it through section 3

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx} % Allows including images
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{tikz}
% \usepackage[capitalize]{cleveref}
\usetikzlibrary{bayesnet}
\setcitestyle{aysep={ }} %Author-Year Separator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

@@

\newcommand{\paren}[1]{\mathopen{}\mathclose\bgroup\left(#1\aftergroup\egroup\right)}
\newcommand{\brock}[1]{\mathopen{}\mathclose\bgroup\left[#1\aftergroup\egroup\right\right]}
\newcommand{\curly}[1]{\mathopen{}\mathclose\bgroup\left\{#1\aftergroup\egroup\right\right\}}
\newcommand{\anglebrackets}[1]{\langle #1 \rangle}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\floor}[1]{\lfloor #1 \rfloor}

@@

\author{First Author         \and
        Second Author         \and
        Third Author%etc.
}

@@

@@

\tim{Is the long citation format required?}
The rest of the paper is organized as follows: related work, including MDL principle discretization \citep{Friedman_1996} and MODL discretization \citet{Boulle_2006} are summarized in \todo{Section 2. replace all of these throughout the paper with cref, ex: $cref{relat_work}$} Preliminaries and notation are covered in Section 3. The proposed single-variable discretization method is derived in Section 4 and is extended to Bayesian networks with multiple continuous variables and combined with Bayesian structure learning in Section 5. The paper concludes with the comparison of the proposed methods against prior work on real-world data in Section 6.

\section{Related Work}
\label{relat_work} % NOTE(tim): I recommend doing \label{sec:relat_work} for sections, eq: for equations, fig: for figures, etc.
\tim{Try to avoid we}
This section reviews MDL principle discretization \citep{Friedman_1996} and MODL discretization \citep{Boulle_2006}. The former is the predominant discretization method for continuous variables in Bayesian networks, and we will compare it to our proposed method in Section 6. The latter is a discretization method for one continuous variable against a discrete target class. Our proposed method generalizes the MODL Bayesian discretization approach to Bayesian networks. The asymptotical equivalence between the MDL and MODL methods were examined in \citep{VL_2000}.

\subsection{MDL Principle Discretization}
\label{MDL}
The MDL principle was first proposed by \cite{MDL_1978}, and states that the best model for a dataset is one which minimizes the amount of information needed to describe it \citep{Grunwald_2009}. MDL methods trade off goodness-of-fit against model complexity to avoid overfitting. In the context of Bayesian networks, \cite{Friedman_1996} applied the MDL principle to determine the optimal number of discretization intervals for continuous variables and optimal the positions of their discretization boundaries. Their approach selects a discretization policy that minimizes the sum of the description length of discretized Bayesian network and the description length of the information necessary for recovering the original continuous values from the discretized data. In the single continuous variable case, the objective function is
\tim{Do you need double vertical bars? If we use single bars it might fit without having to use the small font size.}

\begin{small}
\begin{equation}
\begin{aligned}
\frac{1}{2} \log(N) \left\lbrace  \lVert \Pi_{X_i} \rVert (\lVert X^*_i \rVert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lVert \Pi_{X^*_j} \rVert (\lVert X_j \rVert - 1) \right\rbrace + \log(\lVert {X^*_i} \rVert) \\
 + (N_i-1) H \left( {{||X^*_i|| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right],
\end{aligned}
\end{equation}
\end{small}

\noindent
where $I(\boldsymbol{A},\boldsymbol{B}) = \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$ \tim{consider deleting the right half of equation} is the mutual information between two discrete variables $\boldsymbol{A}$ and $\boldsymbol{B}$, $||\boldsymbol{X}||$ is the cardinality of discrete variable $\boldsymbol{X}$, $X^*_i$ is the discretized version of continuous variable $X_i$ and $H(p) = -p \log(p) - (1-p) \log(1-p)$. The optimal discretization policy can be found by dynamical programming. If there are multiple continuous variables in a Bayesian network, we apply the algorithm on one variable at a time and iterate over all continuous variables. While iterating, only one variable is treated as continuous and other continuous variables are discretized based on initial prediscretization or the discretization result of previous iteration.

For the above discretization processes, the network structure is known in advance. If the network structure is not given, we can alternate between structure learning and discretization learning. We start with some discretization, and learn a network structure given this discretization. Then, we rediscretize based on the learned network. This cycle continues until no improvement in objective function is made.
\subsection{MODL Discretization}
\label{MODL}
MODL discretization \citep{Boulle_2006} is a Bayesian-approach discretization. It discretizes a continuous feature according to a class variable. The best discretization model is found by maximizing the probability $P(\textit{Model} \mid \textit{Data})$ of the model given the data. Using Bayes' rule and since $P(\textit{Data})$ is a constant while varying the discretization model, this is equivalent to maximize:
\begin{align}
P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model}).
\end{align}
Once the prior distribution of the discretization model is fixed, each term in Eq (3) can be evaluated. Usually, $P(\textit{Data} \mid \textit{Model})$ increases as the number of intervals of discretization model increases, since more intervals means that the model contains more information and thus it is easier to reproduce the original data. On the other hand, $P(\textit{Model})$ decreases as the number of intervals increases. Thus, maximizing $P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})$ provides a trade-off to determine number of intervals after discretization.

The MODL method adapts dynamical programming to find the optimal discretization model, and it takes running time $O(N^3 + ||\text{Class}|| \cdot N^2)$, where $N$ is number of data instances. \cite{Lustgarten_2011} provide a different prior to evaluate $P(\text{Model})$ term. This prior could include the assumption of uniform prior probabilities over discretization in MODL as a special case. Additionally, it reduces runing time to $O(||\text{Class}|| \cdot N^2)$.

\section{Preliminaries}
\label{Prelim_Notation}
In this section we provide a brief review of Bayesian networks, including the factorization of joint probability distribution, sampling from a given network, and structure learning. These concepts will be used in later sections. In addition, the formal definition of a discretization policy on a continuous variable is also introduced.

\subsection{Bayesian Network and Structure Learning}
\label{BN_basic}
A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variable $X = \{  X_1, X_2, \cdots, X_n \}$, and whose edges $E$ represents probabilistic dependencies among nodes. The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents in $G$. The second component of $B$, namely $\Theta$, contains a set of parameters that qualify the network. Elements of $\Theta$ take the form $\theta_{x_i \mid \Pi_{x_i}} = P(x_i \mid \Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$ (the set of parents of $X_i$ in $G$). Applying Markov property, we can represent the multivariate joint distribution over $X$ as
\begin{equation}
P_B (X_1 , \cdots, X_n) = \prod_{i=1}^{n} P_B (X_i \mid \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i \mid \Pi_{x_i}}
\end{equation}

For a given Bayesian network, we can generate its data instances by forward sampling, i.e., sampling variables one by one in a topological order \citep[see][chap.~22]{algo_2009}. Given an instance of $X_i$'s parent set, $\Pi_{X_i}$, values of $X_i$ can be sampled according to the conditional probability table $P(X_i \mid \Pi_{X_i})$. Furthermore, this parent-child sampling order can be reversed if we know the marginal probability of child variable $P(X_i)$. By Bayes' rule,
\begin{equation}
P(X_i \mid \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \{ \Pi_{X_i} \}_j) = P(X_i) \cdot P(\Pi_{X_i} \mid X_i),
\end{equation}
where $Pa( X_i)$ is the number of parents of $X_i$, and $\{ \Pi_{X_i} \}_j$ is the $j$th parent of $X_i$. We can first sample $X_i$ by $P(X_i)$, then sample all the parent variables simultaneously by $P(\Pi_{X_i} \mid X_i)$. That is to say, $X_i$ becomes the starting point of sampling.

If we don't know the structure of the Bayesian network in advance, then we need to learn it from data. Roughly speaking, there are three approaches to learn a Bayesian network structure from data \citep[see][chap.~18]{PGM_2009}: constraint-based structure learning, score-based structure learning, and Bayesian model averaging. Here we review the K2 structure learning algorithm \citep{K2}, which is one of the most successful score-based structure learning methods. Similar to most structure learning algorithms, the K2 algorithm requires all variables to be discrete. The score of a learned network is defined as $\prod_i f(X_i, \Pi_{X_i})$, where
\begin{align}
f(X_i, \Pi_{X_i}) = \prod_{j=1}^{||\Pi_{X_i}||} {{(||X_i||-1)!}\over{(N_{ij} + ||X_i||-1)!}} \prod_{k=1}^{||X_i||} \alpha_{ijk}!.
\end{align}
$||X_i||$ is the cardinality of variable $X_i$. $||\Pi_{X_i}||$ is number of all possible instantiations of the parent variables of $X_i$, i.e., $|\Pi_{X_i}| = \prod_{Y \in \Pi_{X_i}} |Y|$. $\alpha_{ijk}$ is the number of instances in a data set that variable $X_i$ is instantiated with its $k$th value, and the parent of $X_i$ are instantiated with the $j$th value of $\Pi_{X_i}$. $N_{ij} = \sum_{k=1}^{|X_i|} \alpha_{ijk}$.

The K2 algorithm searches the network structure with the highest score, which can be interpreted as the most probable network with the given data. In addition, the K2 requires a topological order of variables to be known before the scoring can start. This constraint can prevent cycles from being introduced. The searching for high-score network is an iterative process. There is no way to find the optimal network directly, since there are $2^{O(n^2)}$ possible structures, where $n$ is number of variables. In order to compensate for this, we run the K2 algorithm many times, and each time we start with a different order of variables. The network with the highest score in iterations is the desired one.

\subsection{Discretization Policy}
\label{Disc_p}
A Discretization Policy $M_C = < t_1,t_2, \ldots, t_{k-1} >$ on a continuous variable $C$ is a mapping from $\boldsymbol{R}$ to $\{1,2,3,\ldots,k \}$ such that
\begin{equation}
  M_C (x)=\begin{cases}
    1, & \text{if $x<t_1$}.\\
    i, & \text{if $t_{i-1} \leq x < t_i$}.\\
    k, & \text{if $t_k \leq x$}
  \end{cases}
\end{equation}
That is to say, the policy discretizes the continuous variable $C$ into $k$ intervals. Furthermore, we assume that the discretization edge $t_i$ can only take values on middle points of two consective values of data of $C$. By this assumption, we can obtain an integer representation of $M_C$ as follows:
we sort the data of continuous variable $C$ in an increasing order, $\{ c_1,c_2,\ldots,c_N \}$, and if $t_i = (c_{s_{i}} + c_{s_{i}+1})/2$ for $i=1,2,\ldots,k$, then $M_C = < t_1,t_2, \ldots, t_{k-1} > \equiv [n_1,n_2,..n_k]$, where $n_1 = s_1$ and $n_i = s_{i} - s_{i-1}$ for $i = 2,\ldots,k$.\\

\section{Discretize One Continuous Variable}
\label{one-conti}
In this section, we consider the case that only one variable in a Bayesian network is continuous and the others are discrete. Before we formulate the objective function, we introduce some notations that will make the following calculation easier.
\subsection{Notations}
\label{Notation}
Assume the continuous variable is $X$. It has $n_p$ parent variables, $\Pi_X = \{ P_1, P_2, P_3,\ldots,P_{n_p}\}$. It also has $n_c$ child variables $\{ C_1, C_2,\ldots,C_{n_c}\}$. Each child variable $C_i$ of $X$ has other parent variables, which forms a set $\boldsymbol{S_i}$. Before we do the discretization, we sort the data instances by its attribute of $X$.

@@

\subsection{Approximation}
\label{approx}
Notice that Algorithm1 leads to the non-polynomial runtime $O({v'_{max}}^{n_p} \cdot N^2)$ in the discretization process. Therefore, we have an empirical approximation to the objective function that can reduce runtime and still preserve the quality of discretization. The approximation is as follows: for the dominator of the last factor in Equation 11, we have
\begin{equation}
{{{{n}_i}!}\over{ {{n^{(p)}}_{i,1} !} {n^{(p)}_{i,2} !} \cdots {n^{(p)}_{i,J_P} !}}} \approx \prod_{r=1}^{n_p} { {{{n}_i}!}\over{ {n^{(p_r)}_{i,1} !} {n^{(p_r)}_{i,2} !} \cdots {n^{(p_r)}_{i,J_{p_r}} !}}},
\end{equation}
where $J_{p_r} = || P_r||$ and ${n^{(p_r)}_{i,j} !}$ is the number of instances in $i$th interval with $j$th value of $P_r$. Applying this approximation, the approximated objective function is
\begin{equation}
\begin{aligned}
& \sum_{i=1}^{k-1} - \log(1 - {\exp(- L \cdot {{x_{s_i+1} - x_{s_i}}\over{x_N - x_1}})}) +  \sum_{i=1}^{k} L \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}} +\\
&  \sum_{j=1}^{n_c} \sum_{i=1}^{k}  \sum_{l=1}^{L_j} \left[  \log{{n_{i,l} + J_j - 1}\choose{J_j-1}} + \log \left( { {{n_{i,l}}!}\over{ {n^{(j)}_{i,1,l} !} {n^{(j)}_{i,2,l} !} \cdots {n^{(j)}_{i,J_j,l} !}} } \right) \right] + \\
& \sum_{i=1}^k \left[  \log {{n_{i} + J_P - 1}\choose{J_P-1}} + \sum_{q=1}^{n_p} \log \left( { {{{n}_i}!}\over{ {n^{(p_q)}_{i,1} !} {n^{(p_q)}_{i,2} !} \cdots {n^{(p_q)}_{i,J_{p_q}} !}}} \right) \right].
\end{aligned}
\end{equation}

As we will see in Section 6, the approximated algorithm is more sensitive to the distribution of other variables' values and usually leads to slightly more discretization edges. With the approximation, we can reduce runtime of Algorithm 1 from $O(n_c  {v'_{max}} \cdot N^2 + {v'_{max}}^{n_p} \cdot N^2)$ to $O( {v'_{max}} (n_c + n_p) \cdot N^2)$. The rigorous proof of the approximation have not been proposed.

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx} % Allows including images
\usepackage{epstopdf}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[round]{natbib}
\usepackage{amsfonts} % for mathbb
\usepackage{tikz}
% \usepackage[capitalize]{cleveref}
\usetikzlibrary{bayesnet}
\setcitestyle{aysep={ }} %Author-Year Separator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

@@

\newcommand{\paren}[1]{\mathopen{}\mathclose\bgroup\left(#1\aftergroup\egroup\right)}
\newcommand{\brock}[1]{\mathopen{}\mathclose\bgroup\left[#1\aftergroup\egroup\right]}
\newcommand{\curly}[1]{\mathopen{}\mathclose\bgroup\left\{#1\aftergroup\egroup\right\}}
\newcommand{\anglebrackets}[1]{\langle #1 \rangle}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\floor}[1]{\lfloor #1 \rfloor}

@@

\author{Yi-Chun Chen           \and
        Tim A. Wheeler         \and
        Mykel J. Kochenderfer
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The rest of the paper is organized as follows: related work, including MDL principle discretization \citep{Friedman_1996} and MODL discretization \citet{Boulle_2006} are summarized in Section \ref{sec:related_work}. Preliminaries and notation are covered in Section \ref{sec:preliminaries}. The proposed single-variable discretization method is derived in Section \ref{sec:single_var} and is extended to Bayesian networks with multiple continuous variables and combined with Bayesian structure learning in Section \ref{sec:multi_var}. The paper concludes with the comparison of the proposed methods against prior work on real-world data in Section \ref{sec:experiments}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{sec:related_work}
This section reviews MDL principle discretization \citep{Friedman_1996} and MODL discretization \citep{Boulle_2006}. The former is the predominant discretization method for continuous variables in Bayesian networks, and is compared to our proposed method in Section \ref{sec:experiments}. The latter is a discretization method for one continuous variable against a discrete target class. Our proposed method generalizes the MODL Bayesian discretization approach to Bayesian networks. The asymptotical equivalence between the MDL and MODL methods were examined in \citep{VL_2000}.

\subsection{MDL Principle Discretization}
The MDL principle was first proposed by \cite{MDL_1978}, and states that the best model for a dataset is one which minimizes the amount of information needed to describe it \citep{Grunwald_2009}. MDL methods trade off goodness-of-fit against model complexity to avoid overfitting. In the context of Bayesian networks, \cite{Friedman_1996} applied the MDL principle to determine the optimal number of discretization intervals for continuous variables and optimal the positions of their discretization boundaries. Their approach selects a discretization policy that minimizes the sum of the description length of discretized Bayesian network and the description length of the information necessary for recovering the original continuous values from the discretized data. In the single continuous variable case, the objective function is
\tim{Do you need double vertical bars? If we use single bars it might fit without having to use the small font size.}

\begin{small}
\begin{equation}
\begin{aligned}
\frac{1}{2} \log(N) \left\lbrace  \lVert \Pi_{X_i} \rVert (\lVert X^*_i \rVert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lVert \Pi_{X^*_j} \rVert (\lVert X_j \rVert - 1) \right\rbrace + \log(\lVert {X^*_i} \rVert) \\
 + (N_i-1) H \left( {{\|X^*_i\| - 1}\over{N_i -1}}  \right) -N \cdot \left[ I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j}) \right],
\end{aligned}
\end{equation}
\end{small}

\noindent
where $I(\boldsymbol{A},\boldsymbol{B}) = \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$ \tim{consider deleting the right half of equation} is the mutual information between two discrete variables $\boldsymbol{A}$ and $\boldsymbol{B}$, $\|\boldsymbol{X}\|$ is the cardinality of discrete variable $\boldsymbol{X}$, $X^*_i$ is the discretized version of continuous variable $X_i$ and $H(p) = -p \log(p) - (1-p) \log(1-p)$. The optimal discretization policy is found with dynamical programming. If there are multiple continuous variables in a Bayesian network, we apply the algorithm to one variable at a time and iterate over all continuous variables. While iterating, only one variable is treated as continuous and other continuous variables are discretized based on an initial discretization scheme or the discretization result from a previous iteration.

The network structure is not always known in advance. Bayesian network structure learning can be extended to distributions including continuous variables by alternating between traditional discrete structure learning and optimal discretization. Starting with some preliminary discretization scheme, one applys a structure learning algorithm to identify the locally optimal graph structure. The discretization scheme is refined based on the learned network. The cycle is continued until convergence.

\subsection{MODL Discretization}
\tim{Does MODL stand for something?}
MODL discretization \citep{Boulle_2006} is a Bayesian method for discretizing a continuous feature according to a class variable. Bayesian methods maximize the probability of the model given the data, $P(\textit{Model} \mid \textit{Data})$. An application of Bayes' rule and constant $P(\textit{Data})$ shows that this is equivalent to maximizing:
\begin{align}
P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})\text{,}
\end{align}

\noindent
where $P(\textit{Model})$ is a prior over models and $P(\textit{Data} \mid \textit{Model})$ is the data likelihood.
The data likelihood usually increases with the number of discretization intervals, as more intervals allow for more detailed representations.
The prior usually decreases as the number of discretization intervals increases to favor simpler models and penalize overfitting.
Thus, maximizing $P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})$ requires a trade-off when determining the number of discretization intervals.

The MODL method uses dynamic programming to find the optimal discretization scheme, and has a $\bigo\paren{N^3 + \|\text{Class}\| \cdot N^2}$ runtime, where $N$ is the of number of data instances. \cite{Lustgarten_2011} suggest several formulations for the prior term $P(\text{Model})$\tim{Did I state this correctly?}. If the prior is chosen to reflect uniform prior probability over discretizations the runtime can be reduced to $\bigo\paren{\|\text{Class}\| \cdot N^2}$.\tim{Is a sacrifice made? It is sub-optimal, right?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminaries}
\label{sec:preliminaries}
This section provides a brief overview of Bayesian networks, including the factorization of joint probability distributions, sampling from a Bayesian network, and structure learning. These concepts will be used in later sections. The formal definition of a discretization policy for a continuous variable is also given.

\subsection{Bayesian Networks and Structure Learning}

A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variables $X = \brock{X_1, X_2, \cdots, X_n}$, and whose edges $E$ represent probabilistic dependencies among nodes. The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents $\Pi{X_i}$ in $G$. The conditional relations are parameterized by the elements of $\Theta$, which take the form $\theta_{x_i \mid \Pi_{x_i}} = P(x_i \mid \Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$. The joint distribution over $X$ factors according to the Markov property:
\begin{equation}
P_B (X_1 , \cdots, X_n) = \prod_{i=1}^{n} P_B (X_i \mid \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i \mid \Pi_{x_i}}\text{.}
\end{equation}

Samples can be drawn from the joint distribution represented by a Bayesian network by forward sampling.
Variables are sampled in topological order such that parental instantiations are always known~\citep[see][chap.~22]{algo_2009}.
Furthermore, the parent\rightarrow child sampling order can be reversed if the marginal probability of child variable $P(X_i)$ is known.\tim{Do we use this later?}
By Bayes' rule,

\begin{equation}
P(X_i \mid \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \{ \Pi_{X_i} \}_j) = P(X_i) \cdot P(\Pi_{X_i} \mid X_i),
\end{equation}

\noindent
where $Pa(X_i)$ is the number of parents of $X_i$ and $\brock{\Pi_{X_i}}_j$ is the $j$th parent of $X_i$.
% One can first sample $X_i$ by $P(X_i)$, then sample all the parent variables simultaneously by $P(\Pi_{X_i} \mid X_i)$. That is to say, $X_i$ becomes the starting point of sampling.

It is often necessary to infer the structure of a Bayesian network from data.
Three common approaches to Bayesian network structure learning are constraint-based structure learning, score-based structure learning, and Bayesian model averaging \citep[see][chap.~18]{PGM_2009}. This work uses the K2 structure learning algorithm \citep{K2}, one of the most successful score-based structure learning methods.
Score-based structure learning methods over discrete variables commonly evaluate candidate structures are scored according to their likelihood, $\prod_i f(X_i, \Pi_{X_i})$, where

\begin{align}
f(X_i, \Pi_{X_i}) = \prod_{j=1}^{\|\Pi_{X_i}\|} {{(\|X_i\|-1)!}\over{(N_{ij} + \|X_i\|-1)!}} \prod_{k=1}^{\|X_i\|} \alpha_{ijk}!\text{.}
\end{align}

\noindent
$\|X_i\|$ is the cardinality of $X_i$, $\|\Pi_{X_i}\|$ is number of possible instantiations of the parent variables of $X_i$, $\alpha_{ijk}$ is the number of instances of the $k$th value of $X_i$ and the $j$th parental instantiation, and $N_{ij} = \sum_{k=1}^{|X_i|} \alpha_{ijk}$.\tim{Mykel uses $\alpha_{ij0}$}
The optimal structure is the most probably structure given the data.
The space of acyclic graphcs is superexponential in the number of nodes \citep{PGM_2009} so it is common to rely on heuristic search strategies.

The K2 algorithm assumes a topological ordering of variables and greedily adds parents to nodes to maximally increase the likelihood score.
A fixed ordering ensures acyclicity but does not guarantee a globally optimal network structure.
K2 is typically run multiple times with random topological orderings and the network with the highest likelihood is retained.

\subsection{Discretization Policies}

A discretization policy $M_C = \anglebrackets{t_1 < t_2 < \ldots < t_{k-1}}$ for a continuous variable $C$ is a mapping from $\mathbb{R}$ to $\brock{1,2,3,\ldots,k}$ such that
\begin{equation}
  M_C (x)=\begin{cases}
    1, & \text{if $x<t_1$}.\\
    i, & \text{if $t_{i-1} \leq x < t_i$}\text{.}\\
    k, & \text{if $t_{k-1} \leq x$}
  \end{cases}
\end{equation}

\noindent
The policy discretizes the continuous variable $C$ into $k$ intervals. Furthermore, we restrict the discretization edges $t_\cdot$ to midpoints between two sorted instances of $C$. An integer representation of $M_C$ can be obtained by sorting the samples of $C$ in an increasing order, $\curly{c_1,c_2,\ldots,c_N}$, and if $t_i = (c_{s_{i}} + c_{s_{i}+1})/2$ for $i=1,2,\ldots,k$, then $M_C = \anglebrackets{t_1,t_2, \ldots, t_{k-1}} \equiv \brock{n_1,n_2,..n_k}$, where $n_1 = s_1$ and $n_i = s_{i} - s_{i-1}$ for all $i = 2,\ldots,k$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discretization of a Single Continuous Variable}
\label{sec:single_var}

In this section, we consider the case that only one variable in a Bayesian network is continuous and the others are discrete. Before we formulate the objective function, we introduce some notations that will make the following calculation easier.
\subsection{Notations}
\label{Notation}
Assume the continuous variable is $X$. It has $n_p$ parent variables, $\Pi_X = \{ P_1, P_2, P_3,\ldots,P_{n_p}\}$. It also has $n_c$ child variables $\{ C_1, C_2,\ldots,C_{n_c}\}$. Each child variable $C_i$ of $X$ has other parent variables, which forms a set $\boldsymbol{S_i}$. Before we do the discretization, we sort the data instances by its attribute of $X$.

@@

\subsection{Approximation}
\label{approx}
Notice that Algorithm1 leads to the non-polynomial runtime $O({v'_{max}}^{n_p} \cdot N^2)$ in the discretization process. Therefore, we have an empirical approximation to the objective function that can reduce runtime and still preserve the quality of discretization. The approximation is as follows: for the dominator of the last factor in Equation 11, we have
\begin{equation}
{{{{n}_i}!}\over{ {{n^{(p)}}_{i,1} !} {n^{(p)}_{i,2} !} \cdots {n^{(p)}_{i,J_P} !}}} \approx \prod_{r=1}^{n_p} { {{{n}_i}!}\over{ {n^{(p_r)}_{i,1} !} {n^{(p_r)}_{i,2} !} \cdots {n^{(p_r)}_{i,J_{p_r}} !}}},
\end{equation}
where $J_{p_r} = \| P_r\|$ and ${n^{(p_r)}_{i,j} !}$ is the number of instances in $i$th interval with $j$th value of $P_r$. Applying this approximation, the approximated objective function is
\begin{equation}
\begin{aligned}
& \sum_{i=1}^{k-1} - \log(1 - {\exp(- L \cdot {{x_{s_i+1} - x_{s_i}}\over{x_N - x_1}})}) +  \sum_{i=1}^{k} L \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}} +\\
&  \sum_{j=1}^{n_c} \sum_{i=1}^{k}  \sum_{l=1}^{L_j} \left[  \log{{n_{i,l} + J_j - 1}\choose{J_j-1}} + \log \left( { {{n_{i,l}}!}\over{ {n^{(j)}_{i,1,l} !} {n^{(j)}_{i,2,l} !} \cdots {n^{(j)}_{i,J_j,l} !}} } \right) \right] + \\
& \sum_{i=1}^k \left[  \log {{n_{i} + J_P - 1}\choose{J_P-1}} + \sum_{q=1}^{n_p} \log \left( { {{{n}_i}!}\over{ {n^{(p_q)}_{i,1} !} {n^{(p_q)}_{i,2} !} \cdots {n^{(p_q)}_{i,J_{p_q}} !}}} \right) \right].
\end{aligned}
\end{equation}

As we will see in Section \ref{sec:experiments}, the approximated algorithm is more sensitive to the distribution of other variables' values and usually leads to slightly more discretization edges. With the approximation, we can reduce runtime of Algorithm 1 from $O(n_c  {v'_{max}} \cdot N^2 + {v'_{max}}^{n_p} \cdot N^2)$ to $O( {v'_{max}} (n_c + n_p) \cdot N^2)$. The rigorous proof of the approximation have not been proposed.
